Title

### 2022-05-03 thesis/main.tex
carlos revision 1
change subsection to section and labels
fix section titles

	\chapter{Proposed Models} \label{chapter:models}
	In this chapter, the proposed power, performance and energy models are derived and validated.
	Energy consumption is crucial in high-performance computing (HPC), especially to enable the next exascale generation. Hence, modern systems implement various hardware and software features for power management. Nonetheless, due to numerous different implementations, we can always push the limits of software to achieve the most efficient use of our hardware. To be
	energy efficient, the software relies on dynamic voltage and frequency scaling (DVFS), as well as dynamic power management (DPM). Yet, none have privileged information on the hardware architecture and application behavior, which may lead to energy-inefficient software operation. 
	This study  proposes analytical modeling for architecture and application behavior that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications.
	Additionally, model parameters, such as the level of parallelism and dynamic power, provide insights into how the modeled application consumes energy, which can be helpful for energy-efficient software development and operation.
	This novel analytical model takes the number of active cores, the operating frequencies, and the input size as inputs to provide energy consumption estimation.
	We present the modeling of 13 parallel applications employed to determine energy-optimal configurations for several different input sizes.
	The results show that up to 70\% of energy could be saved in the best scenario compared to the default Linux choice and 14\% on average.
	We also compare the proposed model with standard machine-learning modeling concerning training overhead and accuracy. The results show that our approach generates about 10 times less energy overhead for the same level of accuracy.
	\include{models/models}
	\include{phases/phases}
	%\include{optmizers/optmizers}

@@

	\chapter{Experiments and Results}
	This chapter will describe all the experiments that validate the models constraints.
	\include{experiments/experiments}

>>>>

	\chapter{Application Models} \label{chapter:models}
	In this chapter, the proposed power, performance and energy models are derived and validated.
	Energy consumption is crucial in high-performance computing (HPC), especially to enable the next exascale generation. Hence, modern systems implement various hardware and software features for power management. Nonetheless, due to numerous different implementations, we can always push the limits of software to achieve the most efficient use of our hardware. To be
	energy efficient, the software relies on dynamic voltage and frequency scaling (DVFS), as well as dynamic power management (DPM). Yet, none have privileged information on the hardware architecture and application behavior, which may lead to energy-inefficient software operation. 
	This study  proposes analytical modeling for architecture and application behavior that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications.
	Additionally, model parameters, such as the level of parallelism and dynamic power, provide insights into how the modeled application consumes energy, which can be helpful for energy-efficient software development and operation.
	This novel analytical model takes the number of active cores, the operating frequencies, and the input size as inputs to provide energy consumption estimation.
	We present the modeling of 13 parallel applications employed to determine energy-optimal configurations for several different input sizes.
	The results show that up to 70\% of energy could be saved in the best scenario compared to the default Linux choice and 14\% on average.
	We also compare the proposed model with standard machine-learning modeling concerning training overhead and accuracy. The results show that our approach generates about 10 times less energy overhead for the same level of accuracy.
	\include{models/models}
	%\include{optmizers/optmizers}

	\chapter{Energy optimization using the models}
	\include{experiments/setup}
	\include{optmization/optmization}

	\chapter{Extending the models with phases}
	\include{phases/phases}

@@

	This chapter will describe all the experiments that validate the models constraints.
	\include{framework/validation}

### 2022-05-03 thesis/optmization/optmization.tex
fix section titles

\section{Model Validation} \label{sec:model_validation}
In this section, the models presented in~Sections \ref{sec:power_model} and~\ref{sec:performance_model} were validated with a benchmark specific for multi-core architectures. Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches. We compared against support vector regression (SVR)~\cite{Smola2004ARegression}, decision tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods, such as Gao et al. \cite{Gao2019DendriticPrediction}. However, SVR was chosen as the most representative because it performed best in our tests without aggressive fine-tuning, as shown in \cref{fig:ml_models}.

@@

\section{Fitting the Models} \label{subsec:fitting_the_models}
To find the parameters of  \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$, respectively. The application was executed for each chosen configuration, and the measured energy and time values were collected. For the input size, if we assume that all CPU instructions take approximately the same time to execute, the number of basic operations will be directly correlated with the time. Thus, we can estimate the input size by looking at the execution time, allowing us to divide a large input size into several smaller ones, knowing their relationship, as performed in the work of Oliveira ~\cite{Oliveira2018ApplicationCharacterization}. The unity can also vary depending on the definition. For simplicity, we assign numbers from 1 to 10, increasing the problem linearly, so it is also possible to interpolate any input in between these values.

@@

\section{Measured versus Modeled Energy} \label{sec:measured_versus_modeled_energy}

@@

\section{Overheads on Training} \label{subsec:overhead_on_training}
It is known that machine learning is data-driven; in that sense, the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
To answer that question, the proposed model and the SVR were also trained with a varying number of configurations.
We then compared the MPE and the amount of energy spent to create each model. 
This accuracy-energy trade-off is crucial since building models' energy overhead defeats the primary goal of saving power when running applications.
\enlargethispage{.5cm}

@@

\section{DVFS and DPM Optimization} \label{sec:dvfs_dpm_optminzation}
The effectiveness of the proposed approach during optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the proposed equation. The results were then compared to the Linux default choices for power management.

>>>>

\section{Model validation} \label{sec:model_validation}
In this section, the models presented in~Sections \ref{sec:power_model} and~\ref{sec:performance_model} were validated with a benchmark specific for multi-core architectures. Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches. We compared against support vector regression (SVR)~\cite{Smola2004ARegression}, decision tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods, such as Gao et al. \cite{Gao2019DendriticPrediction}. However, SVR was chosen as the most representative because it performed best in our tests without aggressive fine-tuning, as shown in \cref{fig:ml_models}.

@@

\section{Fitting the models} \label{subsec:fitting_the_models}
To find the parameters of  \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$, respectively. The application was executed for each chosen configuration, and the measured energy and time values were collected. For the input size, if we assume that all CPU instructions take approximately the same time to execute, the number of basic operations will be directly correlated with the time. Thus, we can estimate the input size by looking at the execution time, allowing us to divide a large input size into several smaller ones, knowing their relationship, as performed in the work of Oliveira ~\cite{Oliveira2018ApplicationCharacterization}. The unity can also vary depending on the definition. For simplicity, we assign numbers from 1 to 10, increasing the problem linearly, so it is also possible to interpolate any input in between these values.

@@

\section{Measured versus modeled energy} \label{sec:measured_versus_modeled_energy}

@@

\section{Overheads on training} \label{subsec:overhead_on_training}
It is known that machine learning is data-driven; in that sense, the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
To answer that question, the proposed model and the SVR were also trained with a varying number of configurations.
We then compared the MPE and the amount of energy spent to create each model. 
This accuracy-energy trade-off is crucial since building models' energy overhead defeats the primary goal of saving power when running applications.
\enlargethispage{.5cm}

@@

\section{DVFS and DPM optimization} \label{sec:dvfs_dpm_optminzation}
The effectiveness of the proposed approach during optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the proposed equation. The results were then compared to the Linux default choices for power management.

### 2022-05-03 thesis/pascalanalyzer/pascalanalyzer.tex
fix section titles

\section{Pascal Analyzer} \label{sec:pascal_analyzer}
High-performance computing systems have become increasingly dynamic, complex, and unpredictable. To help build software that uses full-system capabilities, performance measurement and analysis tools exploit extensive execution analysis focusing on single-run results. Despite being effective in identifying performance hotspots and bottlenecks, these tools are not sufficiently suitable to evaluate the overall scalability trends of parallel applications. Either they lack the support for combining data from multiple runs or collect excessive data, causing unnecessary overhead. 
In this work, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies.
%In this work, we present a tool for measuring and comparing several executions of a parallel application automatically with different input arrangements, number of threads, cores, and frequencies to understand its scalability by observing how it uses the available computational resources across configurations. 
Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations. 
%The proposed work covers some gaps and unspecialized features of existing performance analysis tools. 
%The tool features automatic instrumentation and direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use to compare multiple configurations for improving analysis and productivity. 
In order to improve scalability analysis and productivity over the vast spectrum of possible configurations, the proposed tool features automatic instrumentation, direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use.
%In addition, it aims at parallel scalability, instead focus on single-run performance, aggregating a minimal level of intrusion (<1\% on average), which is fundamental for accurately understanding the behavior and scalability trends of the parallel application.
As it aims at accurately understanding scalability trends of parallel applications, detailed single-run performance analyses show minimal intrusion (less than 1\% overhead).

@@

\section{Pascal architecture} \label{sec:pascal_architecture}
The tool we propose to measure, combine and compare multiple runs of a parallel application implements two main concepts to achieve this goal: actuators and sensors.

>>>>

\section{PascalAnalyzer} \label{sec:pascal_analyzer}
High-performance computing systems have become increasingly dynamic, complex, and unpredictable. To help build software that uses full-system capabilities, performance measurement and analysis tools exploit extensive execution analysis focusing on single-run results. Despite being effective in identifying performance hotspots and bottlenecks, these tools are not sufficiently suitable to evaluate the overall scalability trends of parallel applications. Either they lack the support for combining data from multiple runs or collect excessive data, causing unnecessary overhead. 
In this work, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies.
%In this work, we present a tool for measuring and comparing several executions of a parallel application automatically with different input arrangements, number of threads, cores, and frequencies to understand its scalability by observing how it uses the available computational resources across configurations. 
Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations. 
%The proposed work covers some gaps and unspecialized features of existing performance analysis tools. 
%The tool features automatic instrumentation and direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use to compare multiple configurations for improving analysis and productivity. 
In order to improve scalability analysis and productivity over the vast spectrum of possible configurations, the proposed tool features automatic instrumentation, direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use.
%In addition, it aims at parallel scalability, instead focus on single-run performance, aggregating a minimal level of intrusion (<1\% on average), which is fundamental for accurately understanding the behavior and scalability trends of the parallel application.
As it aims at accurately understanding scalability trends of parallel applications, detailed single-run performance analyses show minimal intrusion (less than 1\% overhead).

@@

\section{PascalAnalyzer architecture} \label{sec:pascal_architecture}
The tool we propose to measure, combine and compare multiple runs of a parallel application implements two main concepts to achieve this goal: actuators and sensors.

### 2022-05-03 thesis/phases/phases.tex
fix section titles

\section{Related work: Phases} \label{sec:related_work_phases}
Some related work

@@

\section{Data Collection} \label{sec:data_collection}
Data, power samples, can be gathered by running the application in several possible context configurations. In a system with 32 cores that can run at 13 frequencies, we can have 416 (32*13) possible configurations. Power samples can be taken from dedicated sensors in the IPMI system [REF]. To avoid overload, a sample rate of 0.5 seconds is more than sufficient. Additionally, the behavior of an application can be affected by the workload, so it is also interesting to run applications with different inputs and assess their impact on the collected data [REF victor alex paper].

@@

\section{Results: Phases experiments} \label{sec:results_phases_experiments}

>>>>

\section{Phases related work} \label{sec:related_work_phases}
Some related work

@@

\section{Data collection} \label{sec:data_collection}
Data, power samples, can be gathered by running the application in several possible context configurations. In a system with 32 cores that can run at 13 frequencies, we can have 416 (32*13) possible configurations. Power samples can be taken from dedicated sensors in the IPMI system [REF]. To avoid overload, a sample rate of 0.5 seconds is more than sufficient. Additionally, the behavior of an application can be affected by the workload, so it is also interesting to run applications with different inputs and assess their impact on the collected data [REF victor alex paper].

@@

\section{Phases experiments results} \label{sec:results_phases_experiments}
